---
title: 'Homework: lubridate and purrr'
author: "Luke Govan"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(purrr)
library(dplyr)
```

Exercise 1: Advanced Date Manipulation with lubridate
Question 1: Generate a sequence of dates from January 1, 2015, to December 31, 2025, spaced by every two months. Extract the year, quarter, and ISO week number for each date

```{r}
#Create the sequence of dates
dates <- seq(ymd("2015-01-01"), ymd("2025-12-31"), by = "2 months")

#Show the year quarter and iso_week_number
date_yqI <- data.frame(
  date = dates,
  year = year(dates),
  quarter = quarter(dates),
  iso_week_number = isoweek(dates)
)
print(date_yqI)
```

Exercise 2: Complex Date Arithmetic
Question 2: Given the following dates, compute the difference in months and weeks between each consecutive pair.
```{r}
#Sample dates
sample_dates <- (c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05"))
#Parse 
sample_dates <- as.Date(c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05"))
#Compute difference
difference_result <- data.frame(
  week_date = time_length(diff(sample_dates), "week"),
  month_date = time_length(diff(sample_dates), "month")
)

print(difference_result)
```

Exercise 3:Higher-order Functions with Purrr 
Question 3: Using `map()` and `map_dbl()`, compute the mean, median, and standard deviation for each numeric vector in the list.
```{r}
#Create a numerical list
num_lists <- list(c(4, 16, 25, 36, 49), c(2.3, 5.7, 8.1, 11.4), c(10, 20, 30, 40, 50))

#Use mapdbl to find the mean median and sd
map_dbl(num_lists, mean)
map_dbl(num_lists, median)
map_dbl(num_lists, sd)
```

Exercise 4: Combining lubridate and purr
Question 4: Given a list of mixed date formats, use map() and possibly() from purrr to safely convert them to date format and extract the month name.
```{r}
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")

#parse and then extract the month name
#in order to extract the months actual name we need to use format %B
#because we have an invalid date we need to have a NA when one of the function outputs fails
extracted_month <- possibly(function(x){
  parsed_date <- parse_date_time(x, orders = c("ymd", "dmy", "mdy"), quiet = TRUE)
  format(parsed_date, "%B")
}, otherwise = NA)

#use map to apply the function
monthnames <- map(date_strings, extracted_month)

#print results
resulting_month_name <- data.frame(
  starting_mixed_dates = unlist(date_strings),
  month_name = unlist(monthnames)
)
print(resulting_month_name)

```